---
title: "MCG_Thesis_Section5_analysis"
output: word_document
always_allow_html: true
---
```{r setup, message=FALSE, }
library(dplyr)
library(tibble)
library(afex)
library(emmeans)
library(fastqq)
library(car) #do NOT load this unless all those dplyr::recode were run
library(jtools)
library(stringr)
library(tidyr)
library(knitr)
library(kableExtra)
library(ggthemes)
library(ggplot2)
library(ggbeeswarm)
library(ggpubr)
library(ROCR)

CurrentSourceWD<-dirname(rstudioapi::getSourceEditorContext()$path)
setwd(CurrentSourceWD)

#load all helper functions
setwd('StatsHelperFunctions') #pop into the folder
files.sources = list.files()
sapply(files.sources, source)
setwd(CurrentSourceWD) #go back to yTain folder

aov_get_effect<-function(afexoutput,effect_list){
  #input the table (output of afex)
  #and a vector of the effects (exactly in strings) to extract
  #can also add_atfront --> get one column with custom info, e.g., response variable or model name
  #output the table of...
    #df, MSE, F, p, ges
  d<-cbind(Effect = rownames(afexoutput$anova_table), afexoutput$anova_table)
  d<-d[d$Effect %in% effect_list,] #extract the effect
  
  #rename the columns
  names(d)[names(d)=="num Df"]<-"df_num"
  names(d)[names(d)=="den Df"]<-"df_den"
  names(d)[names(d)=="Pr(>F)"]<-"p"
  
  #finally... remove the rownames
  row.names(d)<-NULL
  
  return(d)
}
MyTwoWayPlot_NoConnect<-function(din,Xvar,Yvar,withinSubjFactor,Ylabel){
  #din is the data frame input; here, be long...
  #Xvar and Yvar and Ylabel should be strings
  
  #the biggest problem: basically incompatible with drawing lines
  
  aus<-ggplot(din,aes_string(x = Xvar, y = Yvar, color = withinSubjFactor)) +
  stat_summary(
    fun.data = mean_cl_boot, size = 1,
    position = position_dodge(0.666) #finally found a line that dodges...
  ) +
  geom_quasirandom(width = .2, stroke = 0, 
                   dodge.width=0.666,
                   alpha=0.8) +
  scale_y_continuous(Ylabel)+theme_classic()
return(aus)
}

```


```{r}
#winning model in fixed-effect comparison
modelname<-c('RW2lrbetaNoBias')

#initialise empty df

## Data preprocessing ~ loading in
omni<-read.csv(sprintf('MCG_OmniTable_PostExclusion_%s.csv',modelname))

omni<-make_lr_logit(omni) #transform raw LR to logit space
  #IMPORTANT: THIS SHOULD NOT BE DONE AGAIN IN FUTURE PROCESSING!!!

omni<-makefactor_vars(omni)
zCovariates<-z_numeric_vars(omni[,c("extID","Age","EducationLevel")])
 #add in medication status (and ALL medication vars) in another analysis, since imbalanced group

#======Create long format over the z scored variables======
#long format for learning rate
longLR<-dplyr::select(omni,extID,group,ANX,ASD,losslrlossvol,losslrwinvol,winlrlossvol,winlrwinvol) %>% 
  pivot_longer(cols = contains('lr'),#,losslrlossvol:winlrwinvol,
               names_to = 'condition',values_to = 'LR') #into long format
#create the within-subject factors
longLR$valence[str_detect(longLR$condition,"winlr")]<-"positive"
longLR$valence[str_detect(longLR$condition,"losslr")]<-"negative"
longLR$block[str_detect(longLR$condition,"winvol")]<-"WinVolatile"
longLR$block[str_detect(longLR$condition,"lossvol")]<-"LossVolatile"
write.csv(longLR,sprintf("MCG_longLR_%s.csv",modelname))
longLR<-z_numeric_vars(longLR) #standardise the learning rates (across all conditions)
longLR<-merge(longLR,
              zCovariates, #list all the controlled variables here
              by="extID")
longLR<- longLR[,!(names(longLR) %in% c("condition"))] #delete the column we no longer need 
longLR<-makefactor_vars(longLR)

#important: tricky line to order the wins before losses when plotting!!
longLR$block<-factor(longLR$block, levels=c("WinVolatile", "LossVolatile"))


#for RLR
theRLR<-dplyr::select(omni,extID,group,ANX,ASD,PositiveRLR,NegativeRLR,AD,CBIT,SW) %>% 
  pivot_longer(cols=c("PositiveRLR","NegativeRLR"),
               names_to = 'valence',values_to = 'RLR')#into long format
#create the within-subject factors
theRLR$valence[str_detect(theRLR$valence,"PositiveRLR")]<-"positive"
theRLR$valence[str_detect(theRLR$valence,"NegativeRLR")]<-"negative"
write.csv(theRLR,sprintf("MCG_long_theRLR_%s.csv",modelname))
theRLR<-z_numeric_vars(theRLR) #do this before adding other already z-scored coolumns
theRLR<-merge(theRLR,
              zCovariates, #add all the control variables here
              by="extID")
theRLR<-makefactor_vars(theRLR)

#====h0 information bias get======
h0 <- afex::aov_car(LR ~ valence*block + Error(extID/valence*block),
                              data = longLR,
                              type = 2) 

#====testing the first two hypotheses=====
h1.2.ctrl <- afex::aov_car(
  LR ~ Age+EducationLevel+ANX*ASD*valence*block + Error(extID/valence*block),
  data = longLR,
  type = 2,
  factorize = FALSE)

#====testing RLR, two hypotheses=====
h3.4.ctrl <- afex::aov_car(
  RLR ~ Age+EducationLevel+ANX*ASD*valence + Error(extID/valence),
                           data = theRLR,
                           type = 2,
                           factorize = FALSE)

#extract the effects of interest
h0df<-aov_get_effect(h0,c("valence:block"))
h1cdf<-aov_get_effect(h1.2.ctrl,c("ANX:valence"))
h2cdf<-aov_get_effect(h1.2.ctrl,c("ASD","ASD:valence"))
h3cdf<-aov_get_effect(h3.4.ctrl,c("ANX","ANX:valence"))
h4cdf<-aov_get_effect(h3.4.ctrl,c("ANX:ASD:valence"))

h3inLRctrl<-aov_get_effect(h1.2.ctrl,c("ANX:valence:block"))




#======doing the analysis over continuous transdiag dimensions=======
h5.ctrl<-lmer(RLR~Age+EducationLevel+valence*(AD+CBIT+SW)+(1 | extID), theRLR)

#for testing if dimensions explain over and above covar of no interest
h5.null<-lmer(RLR~Age+EducationLevel+valence+(1 | extID), theRLR,REML = FALSE)
h5.full<-lmer(RLR~Age+EducationLevel+valence*(AD+CBIT+SW)+(1 | extID), theRLR,REML = FALSE)
h5LRT<-anova(h5.null,h5.full)
```

(Getting the F test results)
```{r}
knitr::kable(h1cdf,digits = 3,caption = "Adjusted for age and education level") %>% 
  kable_classic(full_width = F, html_font = "Cambria")

knitr::kable(h2cdf,digits = 3,caption = "Adjusted for age and education level") %>% 
  kable_classic(full_width = F, html_font = "Cambria")

h3cdf<-h3cdf[order(h3cdf$Effect),]
knitr::kable(h3cdf,digits = 3,caption = "Adjusted for age and education level") %>% 
  #kbl(caption = "Adjusted for age and education level") %>%
  row_spec(which(h3cdf$p<0.025),bold=T) %>% 
  kable_classic(full_width = F, html_font = "Cambria")

knitr::kable(h4cdf,digits = 3,caption = "Adjusted for age and education level") %>% 
  #kbl(caption = "Adjusted for age and education level") %>%
  kable_classic(full_width = F, html_font = "Cambria")

knitr::kable(h3inLRctrl,digits = 3,caption = "Adjusted for age and education level") %>% 
  #kbl(caption = "Adjusted for age and education level") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```
(Getting the linear mixed model results)
```{r}
h5vis<-jtools::plot_summs(h5.ctrl,
                          ci_level = 0.99,
                          inner_ci_level = 0.95,
                          #plot.distributions = TRUE, #not very informative
                          scale = TRUE)

h5vis
```

## Collective model assumption checks
-Just check normality of residuals
```{r}
ResidNorm(h1.2.ctrl)
ResidNorm(h3.4.ctrl)
ResidNorm(h5.ctrl)
```

## Hypothesis 0: induction of information bias
```{r}
knitr::kable(h0df,digits = 3,caption = " ") %>% 
  kable_classic(full_width = F, html_font = "Cambria")

plotIBLT<-emmip(h0,~valence | block, CIs = TRUE)
ValenceBiasTable<-as.data.frame(pairs(emmeans(h0,~valence|block)))
crossBlock<-pairs(emmeans(h0,~block|valence))
LRadjustTable<-as.data.frame(crossBlock)

plotIBLT+theme_classic()+
labs( y = 'Estimated marginal means of \n learning rate (logit space)', x = 'valence',
      title = " ")

knitr::kable(ValenceBiasTable,digits = 3,caption = " ") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
knitr::kable(LRadjustTable,digits = 3,caption = " ") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

```{r}
#Additional checks:
#here be checks for the information bias:
pairs(emmeans(h1.2.ctrl, ~valence | block))
#here be checks for the learning rate regulation by valence
pairs(emmeans(h1.2.ctrl, ~block | valence))
#(basically run the interactions again)
contrast(emmeans(h1.2.ctrl, ~valence*block), interaction = "pairwise") 

```

Additional analysis with ONLY the controls for IBLT
```{r}
#get the data out
longLR_CT<-read.csv(sprintf("MCG_longLR_%s.csv",'RW2lrbetaNoBias'))

longLR_CT<-longLR_CT[longLR_CT$group=="CT",]

CovariatesNoZ<-dplyr::select(omni,extID,Age,EducationLevel)

longLR_CT<-merge(longLR_CT, CovariatesNoZ, 
                 by="extID")
longLR_CT<-z_numeric_vars(longLR_CT) #standardise the learning rates (across all conditions)
longLR_CT<- longLR_CT[,!(names(longLR_CT) %in% c("condition"))] #delete the column we no longer need 
longLR_CT<-makefactor_vars(longLR_CT)
```

```{r}
h0onlyCT <- afex::aov_car(LR ~ valence*block + Error(extID/valence*block),
                              data = longLR_CT) 
knitr::kable(h0onlyCT$anova_table,digits = 3) %>% 
  kable_classic(full_width = F, html_font = "Cambria")

plotIBLT<-emmip(h0onlyCT,~valence | block, CIs = TRUE)
ValenceBiasTable<-as.data.frame(pairs(emmeans(h0onlyCT,~valence|block)))
crossBlock<-pairs(emmeans(h0onlyCT,~block|valence))
LRadjustTable<-as.data.frame(crossBlock)

plotIBLT+theme_classic()+
labs(y = 'Estimated marginal means of \n learning rate (logit space)', x = 'valence',
      title = " ")
knitr::kable(ValenceBiasTable,digits = 3,caption = " ") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
knitr::kable(LRadjustTable,digits = 3,caption = " ") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```




## Follow-up to the unexpected ANX effect directionality

Significant ANOVA result to unpack: H3 possibility 1 ~ main effect of ANX on RLR
```{r}
WithAdjANXmain<-as.data.frame(pairs(emmeans(h3.4.ctrl, ~ANX)))
knitr::kable(WithAdjANXmain,digits = 3,
             caption = "Effect of Anxiety Diagnosis, 
             adjusted for age and education level") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

We hereby go back to the omnibus statistical model over learning rates:

Visualise the IBLT effects:
```{r}
emmeans::emmip(h1.2.ctrl, valence ~ block | ANX,
               #dotarg = list(size = 3),
               CIs = TRUE)+
  ggplot2::theme_classic()+
  labs( y = 'Estimated marginal means of \n learning rate (logit space)', x = 'block')
theEmmipThing<-emmeans::emmip(h1.2.ctrl, valence ~ block | ANX,CIs = TRUE,plotit = FALSE)
datEmmean<-emmeans::emmip(h1.2.ctrl, valence ~ block | ANX) #use this to redo plot by self if necessary

emmip_ggplot(theEmmipThing,
             #dodge=0.2,
             facetlab  = )+
  ggplot2::theme_classic()+
  theme(strip.background = element_blank(), #turn-off facet label for now
            strip.text = element_blank())+
  labs(y = 'Estimated marginal means of \n learning rate (logit space)', x = 'block')+ylim(-0.75,0.50)
```
For the comparisons, stratified by ANX, using the contrasts
```{r}
IntByAnx<-emmeans(h1.2.ctrl,~valence*block|ANX)

IntAll<-emmeans(h1.2.ctrl,~valence*block*ANX)

#as per sequence in the emmeans obj
NegWinVolAnx<-c(1,0,0,0,0,0,0,0)
PosWinVolAnx<-c(0,1,0,0,0,0,0,0)
NegLossVolAnx<-c(0,0,1,0,0,0,0,0)
PosLossVolAnx<-c(0,0,0,1,0,0,0,0)
NegWinVolNoAnx<-c(0,0,0,0,1,0,0,0)
PosWinVolNoAnx<-c(0,0,0,0,0,1,0,0)
NegLossVolNoAnx<-c(0,0,0,0,0,0,1,0)
PosLossVolNoAnx<-c(0,0,0,0,0,0,0,1)
contrast(IntAll,
         method = list("WinInfoBiasAnx" = PosWinVolAnx - NegWinVolAnx,
                       "LossInfoBiasAnx" = NegLossVolAnx - PosLossVolAnx,
                       "NegRLRAnx" = NegLossVolAnx - NegWinVolAnx,
                       "PosRLRAnx" = PosWinVolAnx - PosLossVolAnx,
                       "WinInfoBiasNoAnx" = PosWinVolNoAnx - NegWinVolNoAnx,
                       "LossInfoBiasNoAnx" = NegLossVolNoAnx - PosLossVolNoAnx,
                       "NegRLRNoAnx" = NegLossVolNoAnx - NegWinVolNoAnx,
                       "PosRLRNoAnx" = PosWinVolNoAnx - PosLossVolNoAnx))

```


For split group check of interaction effect:
```{r}
longLR_byANX<-read.csv(sprintf("MCG_longLR_%s.csv",
                               'RW2lrbetaNoBias')) #here, control which model used for

longLR_ANX<-longLR_byANX[longLR_byANX$ANX=="ANX",]
longLR_noANX<-longLR_byANX[longLR_byANX$ANX=="NoANX",]

CovariatesNoZ<-dplyr::select(omni,extID,Age,EducationLevel)


longLR_ANX<-merge(longLR_ANX, 
                  CovariatesNoZ, 
                  by="extID")
longLR_ANX<-z_numeric_vars(longLR_ANX) #standardise the learning rates (across all conditions)
longLR_ANX<- longLR_ANX[,!(names(longLR_ANX) %in% c("condition"))] #delete the column we no longer need 
longLR_ANX<-makefactor_vars(longLR_ANX)


longLR_noANX<-merge(longLR_noANX,
                    CovariatesNoZ,
                    by="extID")
longLR_noANX<-z_numeric_vars(longLR_noANX) #standardise the learning rates (across all conditions)
longLR_noANX<- longLR_noANX[,!(names(longLR_noANX) %in% c("condition"))] #delete the column we no longer need 
longLR_noANX<-makefactor_vars(longLR_noANX)

#runs the pairwise, unpacking the interaction
print('interaction as test of differences: significance of valence*block interactions in ANX vs No-ANX')



IBLT.ANX<-afex::aov_car(LR~Age+EducationLevel+ASD*valence*block + Error(extID/valence*block), #model with all but ANX
                        longLR_ANX,type = 2, factorize = FALSE)
IBLT.NoANX<-afex::aov_car(LR~Age+EducationLevel+ASD*valence*block + Error(extID/valence*block),
                          longLR_noANX,type = 2, factorize = FALSE)
IBLT.ANX
IBLT.NoANX
```


All relevant post-hoc tests:
```{r}
LRaov<-h1.2.ctrl
c0<-contrast(emmeans(LRaov,  ~ANX*valence*block),interaction = "pairwise") #overall three-way interaction
ThreeWayContr<-as.data.frame(c0)

#"information bias change across blocks / two-way interaction, for ANX and non-ANX")
c1<-contrast(emmeans(LRaov, ~valence*ANX*block), interaction = "pairwise", by = "ANX")
ValBlockContr<-as.data.frame(c1)

#"check: simpler differences in information bias, across groups")
c2<-contrast(emmeans(LRaov, ~valence*ANX | block), interaction = "pairwise")
InfoBiasContr<-as.data.frame(c2)

#"check: simpler differences in adjustment of learning rate, across groups")
c3<-contrast(emmeans(LRaov, ~ANX*block | valence), interaction = "pairwise")
LRchangeContr<-as.data.frame(c3)

#"check: simpler differences in learning rate for ANX vs Non-ANX, in each cell")
c4<-contrast(emmeans(LRaov, ~valence*block*ANX), interaction = "pairwise",by=c("valence","block"))
basicLRContr<-as.data.frame(c4)
```

```{r}
knitr::kable(ThreeWayContr,digits = 3,caption = "Contrast for Three-way Interaction of ANX-valence-block") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
knitr::kable(ValBlockContr,digits = 3,caption = "Two-way interactions in ANX groups and in Non-ANX groups") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
knitr::kable(InfoBiasContr,digits = 3,
             caption = "Difference in block-specific information bias, between ANX groups and Non-ANX groups") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
knitr::kable(LRchangeContr,digits = 3,
             caption = "Difference in valence-specific learning rate change, between ANX groups and Non-ANX groups") %>%
  kable_classic(full_width = F, html_font = "Cambria")
knitr::kable(basicLRContr,digits = 3,
             caption = "Simple effects for learning rate differences between ANX groups and Non-ANX groups") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

## Display full ANCOVA table
```{r}
cleanAfexTable<-function(afexoutput){
  d<-cbind(Effect = rownames(afexoutput$anova_table), afexoutput$anova_table)
  
  #rename the columns
  names(d)[names(d)=="num Df"]<-"df_num"
  names(d)[names(d)=="den Df"]<-"df_den"
  names(d)[names(d)=="Pr(>F)"]<-"p"
  row.names(d)<-NULL
  #round digits to three
  
  d$ges<-round(d$ges, digits = 3)
  d$p<-round(d$p, digits = 3)
  
  #get location of any number <0.001 for ges and p
  GesSmall <- d$ges < 0.001
  pSmall <- d$p < 0.001
  
  #make all numbers strings
  d$ges<-as.character(d$ges)
  d$p<-as.character(d$p)
  
  #replace the logged locations with <0.001
  d$ges[GesSmall]<-"<0.001"
  d$p[pSmall]<-"<0.001"
  
  #finally... remove the rownames
  
  return(d)
}

AncovaTable.LR<-cleanAfexTable(h1.2.ctrl)
knitr::kable(AncovaTable.LR,digits = 3,
             caption = "Full ANCOVA table for LR") %>%
  kable_classic(full_width = F, html_font = "Cambria")

AncovaTable.RLR<-cleanAfexTable(h3.4.ctrl)
knitr::kable(AncovaTable.RLR,digits = 3,
             caption = "Full ANCOVA table for RLR") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```


## Exploratory analysis written here for convenience: inverse temperature

```{r}
longB<-Wide2Long2x2(df = omni, theVar = "beta")
  
longB<-merge(longB,dplyr::select(omni,extID,AD,SW,CBIT),by="extID")
longB<-makefactor_vars(longB)
longB$beta<-log(longB$beta) #important: log transform (for normality)
longB<-z_numeric_vars(longB)
```

Test difference in beta
```{r}
beta.ctrl <- afex::aov_car(beta ~ Age+EducationLevel+ANX*ASD*block + Error(extID/block),
                           data = longB,
                           type = 2,
                           factorize = FALSE)
ResidNorm(beta.ctrl)
AncovaTable.beta<-cleanAfexTable(beta.ctrl)
knitr::kable(AncovaTable.beta,digits = 3,
             caption = "Full ANCOVA table for inverse temperature") %>%
  row_spec(which(beta.ctrl$anova_table$`Pr(>F)`<0.05),bold=T) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```
Follow-up to the effect of block
```{r}
pairs(emmeans(beta.ctrl,~block))
emmeans(beta.ctrl,~block)
mean(omni$betawinvol)
mean(omni$betalossvol)
```

for the symptom dimensions
```{r}
h5beta.null<-lmer(beta~Age+EducationLevel+block+(1 | extID), longB,REML = FALSE)
h5beta.full<-lmer(beta~Age+EducationLevel+block*(AD+CBIT+SW)+(1 | extID), longB,REML = FALSE)
h5betatest<-anova(h5beta.null,h5beta.full)
h5betatest
```

## Both-volatile block explorations
```{r}
longBothVol<-dplyr::select(omni,extID,group,
                           ANX,ASD,AD,SW,CBIT,
                           losslrbothvol,winlrbothvol) %>% 
  pivot_longer(cols = contains('lr'),
               names_to = 'condition',values_to = 'LR') #into long format
longBothVol$valence[str_detect(longBothVol$condition,"winlr")]<-"positive"
longBothVol$valence[str_detect(longBothVol$condition,"losslr")]<-"negative"
longBothVol<-z_numeric_vars(longBothVol) #standardise the learning rates (across all conditions)
longBothVol<-merge(longBothVol,
              zCovariates, #list all the controlled variables here
              by="extID")
longBothVol<- longBothVol[,!(names(longBothVol) %in% c("condition"))] #delete the column we no longer need 
longBothVol<-makefactor_vars(longBothVol)

bothvol.ctrl <- afex::aov_car(LR ~ Age+EducationLevel+ANX*ASD*valence + Error(extID/valence),
                           data = longBothVol,
                           type = 2,
                           factorize = FALSE)
ResidNorm(bothvol.ctrl)
bothvol.ctrl
pairs(emmeans(bothvol.ctrl,~valence))


omnilogged<-omni
omnilogged$betabothvol<-log(omnilogged$betabothvol)
bothvol.beta <- afex::aov_car(betabothvol ~ Age+EducationLevel+ANX*ASD + Error(extID),
                           data = omnilogged,
                           type = 2,
                           factorize = FALSE)
ResidNorm(bothvol.beta)
bothvol.beta


bothvolLRdf<-cleanAfexTable(bothvol.ctrl)
knitr::kable(bothvolLRdf,digits = 3,
             caption = " ") %>%
  row_spec(which(bothvol.ctrl$anova_table$`Pr(>F)`<0.05),bold=T) %>% 
  kable_classic(full_width = F, html_font = "Cambria")

```

```{r}
#for testing if dimensions explain over and above covar of no interest
h5bothvol.null<-lmer(LR~Age+EducationLevel+valence+(1 | extID), longBothVol,REML = FALSE)
h5bothvol.full<-lmer(LR~Age+EducationLevel+valence*(AD+CBIT+SW)+(1 | extID), longBothVol,REML = FALSE)
h5bothvoltest<-anova(h5bothvol.null,h5bothvol.full)
h5bothvoltest
```


## Explore the role of AD dimension in conjunction with diagnostic categories
```{r}
afex::aov_car(RLR ~ Age+EducationLevel+AD*ASD*valence + Error(extID/valence),
                           data = theRLR,
                           type = 2,
                           factorize = FALSE)
withADcovar<-afex::aov_car(RLR ~ Age+EducationLevel+ANX*ASD*valence + AD + Error(extID/valence),
                           data = theRLR,
                           type = 2,
                           factorize = FALSE)
withADcovar
pairs(emmeans(withADcovar,~ANX))
```
